TacticToe: Learning to Prove with Tactics

### 2018-02-19 tactictoe_jv.tex
Finish sec 4

\subsection{ATP Integration}\label{sec:atp}
General-purpose proof automation mechanisms which combine proof translation to
ATPs with machine learning (``hammers'') have become quite successful in
enhancing the automation level in proof assistants~\cite{hammers4qed}.
As external automated reasoning techniques sometimes outperform the combined
power of tactics, we would like to combine the \tactictoe search with
general purpose provers such as the ones found in \holyhammer for
\holfour~\cite{tgck-cpp15}.

Such a prover already exists in \holfour and is called \metis. It is already
recognized by the tactic selection mechanism in \tactictoe and can have its
premises predicted dynamically when in an abstracted form. Nevertheless,
we think that the performance of \tactictoe can be boosted by giving the ATP
\metis a special status. This arrangement consists of always predicting
premises for \metis, giving it a slightly higher timeout and
trying it first on each open goal. These modifications only induce an
linear overhead on the number nodes as a general purpose prover does not create
goal obligations.

The integration of an external prover is also possible. And
\eprover~\cite{eprover} is the one we experimented with.
To this end, we created a tactic which expects a number of premises as an
argument
and calls \eprover on the translated first order problem. We then extract the
premises that were required in the \eprover
proof and reconstruct it inside \holfour with \metis. Giving a special status
to external
provers is essential as external ATP calls do not appear in human
proof scripts. For external ATPs, we use an even higher timeout (5 seconds) and
a larger number of predicted premises (128 for \eprover) and also try them
first on open
goals. Since calls to ATPs are
computationally expensive, they are run in parallel and asynchronously. This
avoids slowing down \tactictoe's search loop. The number of asynchronous calls
to external provers that can be executed at the same time is limited by the
number of cores available to the user.

@@

Machine learning has also been used to advise the best library lemmas for new
ITP goals.
This can be done either in an interactive way, when the user completes the
proof based on the recommended lemmas, as in the Mizar proof advisor
\textsc{MizAR}~\cite{Urb04-MPTP0,mizAR40}, or attempted fully automatically, where such lemma
selection is handed over to the ATP component of a \emph{hammer}
system~\cite{hammers4qed,tgck-cpp15,holyhammer,BlanchetteGKKU16}.

>>>>

\subsection{ATP Integration}\label{sec:atp}
General-purpose proof automation mechanisms which combine proof translation to
ATPs with machine learning (``hammers'') have become quite successful in
enhancing the automation level in proof assistants~\cite{hammers4qed}.
Since external automated reasoning techniques sometimes outperform the combined
power of tactics, we would like to combine the \tactictoe search with
general purpose provers such as the ones found in \holyhammer for
\holfour~\cite{tgck-cpp15}.

\holfour already integrates a superposition-based prover \metis for this purpose. It is already
recognized by the tactic selection mechanism in \tactictoe and its
premises can be predicted dynamically when in an abstracted form. Nevertheless,
we think that the performance of \tactictoe can be boosted by giving the ATP
\metis a special status. This arrangement consists of always predicting
premises for \metis, giving it a slightly higher timeout and
trying it first on each open goal. These modifications only induce an
linear overhead on the number nodes as \metis does not create
new goals.

The integration of an external prover is also possible. For this purpose we experiment
with 
\eprover~\cite{eprover}.
Namely, we developed a tactic which expects a list of premises as an
argument
and calls \eprover on the goal together with the premises translated to a first-order problem. If \eprover
succeeds on the problem, the premises used in the external proof are used to
reconstruct the proof inside \holfour with \metis. Giving a special status
to external
provers is essential as external ATP calls do not appear in human
proof scripts. For external ATPs, we use an even higher timeout (5 seconds) and
a larger number of predicted premises (128 for \eprover) and also try them
first on open
goals. Since calls to ATPs are
computationally expensive, they are run in parallel and asynchronously. This
avoids slowing down \tactictoe's search loop. The number of asynchronous calls
to external provers that can be executed at the same time is limited by the
number of cores available to the user.

@@

Machine learning has also been used to advise the best library lemmas for new
ITP goals.
This can be done either in an interactive way, when the user completes the
proof based on the recommended lemmas, as in the Mizar proof advisor
\textsc{MizAR}~\cite{Urb04-MPTP0,mizAR40}, or attempted fully automatically, where such lemma
selection is handed over to the ATP component of a \emph{hammer}
system, as in the hammers for \holfour~\cite{tgck-cpp15}, \hollight~\cite{holyhammer} and \isabelle~\cite{BlanchetteGKKU16}.
